PMC Survey Analysis Part 2

### 2022-01-12 analysis_code/analysis2.Rmd
fixed minor issue with journal name mismatches betwen scimago and nlm

library("wordcloud")
library("vioplot")

@@

@@

nlmcat <- readLines("../data/nlmcatalog_result.txt")
journaltitle <- nlmcat[grep("Title\\(s\\):",nlmcat)]
journaltitle <- gsub("Title\\(s\\): ","",journaltitle)
journaltitle <- gsub("\\.$","",journaltitle)
journaltitle <- toupper(journaltitle) 
journalabbrev <- nlmcat[grep("Title Abbreviation:",nlmcat)]
journalabbrev <- sapply(strsplit(journalabbrev,":"),"[[",2)
journalabbrev <- gsub(" $","",journalabbrev)
journalabbrev <- gsub("^ ","",journalabbrev)
jdf <- data.frame(journalabbrev,journaltitle)

mjdf <- merge(jdf,jmetrics,by.x="journaltitle",by.y="Title")

xm <- merge(x,mjdf,by.x="Journal",by.y="journalabbrev",all.x = TRUE)

@@

@@

@@

xmm <-xm[,c("Journal","scores")]
tab <- table(xmm$Journal)
tab <- tab[which(tab>=5)]
xmm <- xmm[which(xmm$Journal %in% names(tab)),]
dim(xmm)
jres <- aggregate(. ~ Journal, xmm ,mean)
jsd <- aggregate(. ~ Journal, xmm ,sd)
jres$sd <- jsd$scores
jres <- jres[order(jres$scores),]
dim(jres)
head(jres)

@@

j <- unique(xmm$Journal)
jscores <- sapply(j,function(jj) {
 xmm[which(xmm$Journal == jj ),2]  
})

@@

jjres <- merge(jres,mjdf,by.x="Journal",by.y="journalabbrev",all.x = TRUE)

mylm2 <- lm (jjres$scores ~ jjres$SJR)
plot(jjres$SJR,jjres$scores,col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")

png("images/mean_sjr1.png",width=350,height=300)
par(mar=c(5,5,3,1))
plot(jjres$SJR,jjres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

cor.test(jjres$scores,jjres$SJR,method="pearson")

cor.test(jjres$scores,jjres$SJR,method="spearman")

pdf("images/mean_sjr1.pdf",width=3.5,height=3)
par(mar=c(5,5,3,1))
plot(jjres$SJR,jjres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

@@

@@

>>>>

library("wordcloud")
library("vioplot")
library("ggplot2")
library("ggExtra")
library("europepmc")

@@

There is a problem that the journal names from NLM/PUBMED CENTRAL do not exactly match those given by SciMago.
This results in some journals being dropped from the SJR correlation analysis.
To avoid this, I will use the ISSN numbers which should be unique and non-ambiguous identifiers.
This is obtained with a query of Europe PMC with the europepmc package.

@@

getissn <- function(pmc){
  Sys.sleep(1)
  pmcres <- europepmc::epmc_search(pmc)
  pmcissn <- pmcres$journalIssn
  pmcissn <- gsub("-","",pmcissn)
  return(pmcissn)
}

issn <- sapply(x$Pubmed.Central.ID,getissn)
issn2 <- lapply(issn,function(x) { unlist(strsplit(x,"\\; ")) } )
l <- lapply(issn2,length)
l <- unlist(l)
lnames <- names(l)
pmcnames <- unlist(sapply(1:length(l), function(i) { rep(lnames[i],l[i]) }  ))
issn2 <- unname(unlist(issn2))
issn2 <- gsub("-","",issn2)
pmcissn <- data.frame(pmcnames,issn2)
xissn <- merge(x,pmcissn,by.x="Pubmed.Central.ID",by.y="pmcnames")

xm <- merge(xissn,jmetrics,by.x="issn2",by.y="Issn")

@@

Here I will be making a scatter plot of the analysis scores vs the SJR.

I will also attempt to make the same scatterplot with marginal histograms according to (https://stackoverflow.com/questions/8545035/scatterplot-with-marginal-histograms-in-ggplot2)

@@

df <- data.frame("SJR" = xm$SJR, "score" = xm$scores )
p <- ggplot(df, aes(SJR, score)) + 
  geom_point(alpha=0.1, size = 3) + 
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")

pdf("images/score_sjr2.pdf",width=4,height=3)
p <- ggplot(df, aes(SJR, score)) + 
  geom_point(alpha=0.1, size = 3) + 
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")
dev.off()

@@

x2 <-x[,c("Journal","scores")]
tab <- table(x2$Journal)
tab <- tab[which(tab>=5)]
x2 <- x2[which(x2$Journal %in% names(tab)),]
dim(x2)
jres <- aggregate(scores ~ Journal, x2 ,mean)
jsd <- aggregate(scores ~ Journal, x2 ,sd)
jres$sd <- jsd$scores
jres <- jres[order(jres$scores),]
dim(jres)
head(jres)

@@

j <- unique(x2$Journal)
jscores <- sapply(j,function(jj) {
 x2[which(x2$Journal == jj ),2]  
})

@@

x2 <-xm[,c("Journal","SJR","scores")]
tab <- table(x2$Journal)
tab <- tab[which(tab>=5)]
x2 <- x2[which(x2$Journal %in% names(tab)),]
dim(x2)
jres <- aggregate(scores ~ Journal, x2 ,mean)
jsjr <- aggregate(SJR ~ Journal, x2 ,mean)
jres$SJR <- jsjr$SJR
jres <- jres[order(jres$scores),]
dim(jres)
head(jres)

mylm2 <- lm (jres$scores ~ jres$SJR)
plot(jres$SJR,jres$scores,col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")

png("images/mean_sjr1.png",width=350,height=300)
par(mar=c(5,5,3,1))
plot(jres$SJR,jres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

cor.test(jres$scores,jres$SJR,method="pearson")

cor.test(jres$scores,jres$SJR,method="spearman")

pdf("images/mean_sjr1.pdf",width=3.5,height=3)
par(mar=c(5,5,3,1))
plot(jres$SJR,jres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

df <- data.frame("SJR" = jres$SJR, "mean score" = jres$scores )
p <- ggplot(df, aes(SJR, mean.score)) +
  geom_point(alpha=0.3, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")

pdf("images/mean_sjr2.pdf",width=4,height=3)
p <- ggplot(df, aes(SJR, mean.score)) +
  geom_point(alpha=0.3, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")
dev.off()

@@

cor.test(xm$scores , log2(xm$NumCites+1), method="pearson")

cor.test(xm$scores , log2(xm$NumCites+1), method="spearman")

@@

df <- data.frame("log2.citations" = log2(xm$NumCites+1), "score" = xm$scores )

p <- ggplot(df, aes(log2.citations, score)) +
  geom_point(alpha=0.1, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")

pdf("images/cites2.pdf",width=4,height=3)
p <- ggplot(df, aes(log2.citations,score)) +
  geom_point(alpha=0.1, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")
dev.off()
